---
title: "Claire Cleans Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r include=FALSE}
library(roxygen2)
```

# Get to know data
```{r setup}
library(devtools)
devtools::install_github("cecusack/clairecleans")
library(dplyr)

#> to load datasets in this package
data(dat, package="clairecleans")
data(fake_data, package="clairecleans")

#> quick view data
head(dat)
head(fake_data)
```
# fixheight

`fixheight()` takes two arguments, the name of the dataframe and the name of the variable containing heights in quotes.    

When you call the `fixheight()` function, you will receive a return message, a dataframe with flagged heights only, and a dataframe with updated heights.

## Example 1

```{r}
#> Get a sense of what height looks like
table(dat$height, useNA = "ifany") 
#> a mess
#> to only see unique values call line below
#> unique(dat$height)

fixheight(dat, "height")
#> returns 2 columns, original height and new height. dataframe in global environment should have one extra column that it did prior to running this line
#> you'll also see in your global environment a new dataframe called heightflag. This will include rows with heights that should be looked into. 
```

### Flagged heights
```{r}
heightflag %>% select(height_original, height) 
#> one person said their height was 30 inches and one said it was 1234 inches. Those are implausible values

#> example of how to make them NA
#> not run
#> dat[which(dat$height_original=="30"), "height"] <- NA
#> dat[which(dat$height_original=="1234"), "height"] <- NA

#> example of how to make all in height flagged NA in height cols, where participant id is named "id"
dat[(dat$id %in% heightflag$id), "height"] <- NA
#> check
dat %>% select(height_original, height)
#> you'll notice ID 16 and 17 now have NA in height col

#> after you've addressed heights with problematic values, you can remove the heightflag object from your global environment
#> not run
#> rm(heightflag)
```

## Example 2
```{r}
head(fake_data) # height is var demo_9

#> what does height look like 
table(fake_data$demo_9, useNA = "ifany")

#> use fixheight function
fixheight(fake_data, "demo_9")  
#> scroll through. look fine
#> Notice: there is no dataframe titled heightflag with this example because no observations were identified as implausible

#> check values
psych::describe(fake_data$height) 
#> min = 60, max = 69. mean = 63.74. none of the vals are implausible
```

# fixweight

This function operates with the same conventions as fixheight. 
```{r}
#> see responses
table(dat$weight)

fixweight(dat, "weight")

#> examine rows where weight was flagged
weightflag %>% select(weight_original, weight) # "seventy seven"

#> fix this observation manually
dat[which(dat$weight_original=="seventy seven"), which(colnames(dat)=="weight")] <- 77

#> check again
#> not run
#> dat %>% select(weight_original, weight)

psych::describe(dat$weight) # mean = 147.76, min = 77, max = 274

#> after weights checked, you can clean weightflag from environment
#> rm(weightflag)
```

## Metric units

Change height from inches to centimeters and weight in pounds to kilograms.
```{r}
#> change height from inches to cm
#> not run
#> dat$height=conv_unit(dat$height, from="inch", to="cm")

#> change weight from pounds to kilograms
#> not run
#> dat$weight=conv_unit(dat$weight, from="lbs", to="kg")
```

## Calculate BMI

if you've fixed height and weight, and want to calculate BMI.
```{r}
#> for imperial units inches and pounds
dat$bmi=(dat$weight*703)/(dat$height^2)
#> dat %>% select(height, weight, bmi)

#> for metric units kg and cm
#> not run
#> dat$bmi=(dat$weight/(test$height^2))*10000
#> dat %>% select(height, weight, bmi)
```

# check_measure

Returns quick descriptives (mean, standard deviation, max, min) and a dataframe with columns min_check and max_check.    

The string "Pass" signifies all observations are within range. The string "CHECK" indicates at least one observation is below the minimum or above the maximum.     

Also returns a message listing variables that did fall outside of possible values for that item.    

If multiple items have the same range, you can concatenate a list using examples 2-4 below.
```{r}
#> example of checking one item
check_measure(fake_data$edeq6_1, 0, 6)

#> example checking multiple items using grep()
check_measure(fake_data[,grep('bdi_[0-9]{1,2}_clinical', colnames(fake_data))],0,6)

#> example checking multiple items using column numbers
check_measure(fake_data[,c(41,48)], 0,3)

#> example checking multiple items using c(which(colnames(dat)=="var"), which(colnames(dat)=="var2"))
check_measure(fake_data[,c(which(colnames(fake_data)=="edeq6_1"), which(colnames(fake_data)=="ede6_6"))], 0,6)

```

